Taking Linear Logic Apart

### 2018-10-26 doc/paper/main.tex
Rewrote proofs and theory of TLLA 2018

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathtools}
\def\Coloneqq{\mathbin{::=}}
\def\coloneqq{\mathbin{:=}}
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev,capitalize]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{definition}[lemma]{Definition}
\input{preamble-typing}
\usepackage[sectionbib,sort&compress]{natbib}

@@

        \cpRedKappaTens1
        & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
        & \Longrightarrow \;
        & \tm{\cpSend{y}{z}{\cpCut{x}{P}{R}}{Q}}
          \quad \text{if} \; \notFreeIn{x}{Q}
        \cpRedKappaTens2
        & \tm{\cpCut{x}{\cpSend{y}{z}{P}{Q}}{R}}
        & \Longrightarrow \;
        & \tm{\cpSend{y}{z}{P}{\cpCut{x}{Q}{R}}}
          \quad \text{if} \; \notFreeIn{x}{P}
        \cpRedKappaParr
        & \tm{\cpCut{x}{\cpRecv{y}{z}{P}}{R}}
        & \Longrightarrow \;
        & \tm{\cpRecv{y}{z}{\cpCut{x}{P}{R}}}
        \cpRedKappaBot
        & \tm{\cpCut{x}{\cpWait{y}{P}}{R}}
        & \Longrightarrow \;
        & \tm{\cpWait{y}{\cpCut{x}{P}{R}}}
        \cpRedKappaPlus1
        & \tm{\cpCut{x}{\cpInl{y}{P}}{R}}
        & \Longrightarrow \;
        & \tm{\cpInl{y}{\cpCut{x}{P}{R}}}
        \cpRedKappaPlus2
        &\tm{\cpCut{x}{\cpInr{y}{P}}{R}}
        & \Longrightarrow \;
        & \tm{\cpInr{y}{\cpCut{x}{P}{R}}}
        \cpRedKappaWith
        & \tm{\cpCut{x}{\cpCase{y}{P}{Q}}{R}}
        & \Longrightarrow \;
        & \tm{\cpCase{y}{\cpCut{x}{P}{R}}{\cpCut{x}{Q}{R}}}
        \cpRedKappaTop
        & \tm{\cpCut{x}{\cpAbsurd{y}}{R}}
        & \Longrightarrow \;
        & \tm{\cpAbsurd{y}}

@@

Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function \ie $\ty{(A^\bot)^\bot} = \ty{A}$.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
An environment associates channels with types. Names in environments must be unique, and two environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma, \Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} \Coloneqq \ty{\emptycontext} \mid \ty{\Gamma}, \tmty{x}{A}
  \)
\end{definition}\noindent
A typing judgement associates a process with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}, \dots, \tmty{x_n}{A_n}}$ denotes that the process \tm{P} communicates along channels $\tm{x_1}$, \dots, $\tm{x_n}$ following protocols $\ty{A_1}$, \dots, $\ty{A_n}$. Typing judgements are derived using rules below. 
  \begin{description}
	  \item[Structural rules]
  		\begin{gather*}\cpInfAx  \qquad  \cpInfCut   \end{gather*}
	  \item[Logical rules]
		  \begin{gather*}
				\cpInfTens \qquad 
				\cpInfParr \\
				\cpInfBot \qquad 
				\cpInfOne \\
				\cpInfPlus1 \qquad 
				\cpInfPlus2 \\
				\cpInfWith \qquad
				\cpInfNil \qquad 
				\cpInfTop   
			\end{gather*}
	\end{description}
\end{definition}

\subsection{Metatheory}
\cp enjoys subject reduction, termination, and progress~\citep{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, measured in the size of the cut term. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$. If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
  \item
    if either side of the cut is itself a cut, we recursively eliminate the cut;
  \item
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
  \item
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}
\end{proof}

@@

        \hccpEquivMixHalt1

@@

        \hccpEquivMixComm

@@

        \hccpEquivMixAss1

@@

        \hccpEquivNewComm

@@

        \hccpEquivScopeExt1

@@

      \hccpRedAxCut1

@@

      \hccpRedBetaTensParr

@@

      \hccpRedBetaOneBot

@@

      \hccpRedBetaPlusWith1

@@

      \hccpRedBetaPlusWith2

@@

      \hccpRedKappaNuTens
      & \tm{\piNew{x}{\piBoundSend{y}{z}{P}}}
      & \tm{\piBoundSend{y}{z}{\piNew{x}{P}}}
      \hccpRedKappaNuParr
      & \tm{\piNew{x}{\piRecv{y}{z}{P}}}
      & \tm{\piRecv{y}{z}{\piNew{x}{P}}}
      \hccpRedKappaNuOne
      & \tm{\piNew{x}{\piBoundSend{y}{}{P}}}
      & \tm{\piBoundSend{y}{}{\piNew{x}{P}}}
      \hccpRedKappaNuBot
      & \tm{\piNew{x}{\piRecv{y}{}{P}}}
      & \tm{\piRecv{y}{}{\piNew{x}{P}}}
      \hccpRedKappaNuPlus1
      & \tm{\piNew{x}{\cpInl{y}{P}}}
      & \tm{\cpInl{y}{\piNew{x}{P}}}
      \hccpRedKappaNuPlus2
      & \tm{\piNew{x}{\cpInr{y}{P}}}
      & \tm{\cpInr{y}{\piNew{x}{P}}}
      \hccpRedKappaNuWith
      & \tm{\piNew{x}{\cpCase{y}{P}{Q}}}
      & \tm{\cpCase{y}{\piNew{x}{P}}{\piNew{x}{Q}}}
      \hccpRedKappaNuTop
      & \tm{\piNew{x}{\cpAbsurd{y}}}
      & \tm{\cpAbsurd{y}}
      \hccpRedKappaMixTens
      & \tm{(\piPar{\piBoundSend{y}{z}{P}}{R})}
      & \tm{\piBoundSend{y}{z}{(\piPar{P}{R})}}
      \hccpRedKappaMixParr
      & \tm{(\piPar{\piRecv{y}{z}{P}}{R})}
      & \tm{\piRecv{y}{z}{(\piPar{P}{R})}}
      \hccpRedKappaMixOne
      & \tm{(\piPar{\piBoundSend{y}{}{P}}{R})}
      & \tm{\piBoundSend{y}{}{(\piPar{P}{R})}}
      \hccpRedKappaMixBot
      & \tm{(\piPar{\piRecv{y}{}{P}}{R})}
      & \tm{\piRecv{y}{}{(\piPar{P}{R})}}
      \hccpRedKappaMixPlus1
      & \tm{(\piPar{\cpInl{y}{P}}{R})}
      & \tm{\cpInl{y}{(\piPar{P}{R})}}
      \hccpRedKappaMixPlus2
      & \tm{(\piPar{\cpInr{y}{P}}{R})}
      & \tm{\cpInr{y}{(\piPar{P}{R})}}
      \hccpRedKappaMixWith
      & \tm{(\piPar{\cpCase{y}{P}{Q}}{R})}
      & \tm{\cpCase{y}{(\piPar{P}{R})}{(\piPar{Q}{R})}}
      \hccpRedKappaMixTop
      & \tm{(\piPar{\cpAbsurd{y}}{R})}
      & \tm{\cpAbsurd{y}}

@@

        \SYM{\hccpRedGammaNew}

@@

        \SYM{\hccpRedGammaMix}

@@

        \SYM{\hccpRedGammaEquiv}

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \begin{description}
    \item[Structural rules]
    \begin{gather*}
			\hccpInfAx \qquad
			\hccpInfCycle \\
	    \hccpInfMix  \qquad
	    \hccpInfHalt
    \end{gather*}
    \item[Logical rules]
    \begin{gather*}
    	\hccpInfBoundTens \qquad \hccpInfParr \\
			\hccpInfOne  \qquad    \hccpInfBot \\  
			\hccpInfPlus1 \qquad    \hccpInfPlus2 \\
			\hccpInfWith \\    
			\hccpInfNil   \qquad    \hccpInfTop
    \end{gather*}
  \end{description}
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
\end{definition}\noindent

\subsection{Metatheory}
\hcp enjoys subject reduction, termination, and progress.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$. If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}

@@

    \pvar{$\rho_3$}\UIC{$\seq{\mathcal{G}_1 \hsep {B}, \Gamma}$}
    \pvar{$\rho_4$}\UIC{$\seq{\mathcal{G}_2}$}
    \BIC{$\seq{\mathcal{G}_1 \hsep \mathcal{G}_2 \hsep {B}, \Gamma}$}
    \SYM{\with}
    \BIC{$\seq{\mathcal{G}_1 \hsep \mathcal{G}_2 \hsep {A \with B}, \Gamma}$}
      \pvar{$\rho_3$}\UIC{$\seq{\mathcal{G}_1 \hsep {B}, \Gamma}$}
      \SYM{\with}
      \BIC{$\seq{\mathcal{G}_1 \hsep {A \with B}, \Gamma}$}
      \BIC{$\seq{\mathcal{G}_1 \hsep \mathcal{G}_2 \hsep {A \with B}, \Gamma}$}

@@

Disentanglement is terminating, and confluent up to the associativity and commutativity of mixes.
\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If there exists a derivation $\rho$ of $\seq{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp, then there exist derivations $\rho_1, \dots, \rho_n$ of $\seq{\Gamma_1}, \dots, \seq{\Gamma_n}$ in \cp such that
  \[
  \rho
  \moveMixDown^\star
  \begin{prooftree*}
    \AXC{}\noLine\UIC{$\mtf{\rho_1}$}
    \AXC{}\noLine\UIC{$\dots$}
    \AXC{}\noLine\UIC{$\mtf{\rho_n}$}
    \NOM{H-Mix}
    \TIC{$\seq{\Gamma_1 \hsep \dots \hsep \Gamma_n}$}
  \end{prooftree*}
  \]
\end{lemma} 
\begin{proof}
  We repeatedly apply the $\moveMixDown$-rules to the derivation $\rho$ to move the mixes downwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck under a cut, it forms a \cp cut;
  \item
    if a mix gets stuck under a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the top.
  \end{enumerate*}
  All applications of $(\one)$ are followed by an application of \textsc{H-Halt}, forming a \cp $(\one)$.
\end{proof}
An environment can be internalised as a type by collapsing it as a series of pars.
\begin{definition}\label{def:bigparr}
  \[
  \begin{array}{lcll}
    \ty{\bigparr(\emptycontext)}
    & = & \ty{\bot}
    \\
    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} 
     & = & \ty{A_1 \parr \dots \parr A_n}
     & \text{ if } n \geq 1
    \\
  \end{array}
  \]
\end{definition}\noindent
\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma} 
\begin{proof}
  By repeated application of $(\parr)$.
\end{proof}\noindent
Furthermore, a hyper-environment can be internalised as a type by collapsing it as a series of tensors, where each constituent environment is internalised using $\ty{\bigparr}$. The empty hyper-environment $\ty{\emptyhypercontext}$ is internalised as the unit of tensor.
\begin{definition}\label{def:bigtens}
  \[
  \begin{array}{lcll}
    \ty{\bigtens(\emptyhypercontext)}
    & = & \ty{\one}
    \\
    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
    & \text{ if } n \geq 1
  \end{array}
  \]
\end{definition}\noindent

>>>>

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amsthm}
\usepackage{amssymb}
\def\Coloneqq{\mathbin{::=}}
\def\coloneqq{\mathbin{:=}}
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev,capitalize]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{definition}[lemma]{Definition}
\input{preamble-typing}
\usepackage[sectionbib,sort&compress]{natbib}

@@

@@

Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function \ie $\ty{(A^\bot)^\bot} = \ty{A}$.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
An environment associates channels with types. Names in environments must be unique, and two environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma, \Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} \Coloneqq \ty{\emptycontext} \mid \ty{\Gamma}, \tmty{x}{A}
  \)
\end{definition}\noindent
A typing judgement associates a process with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}, \dots, \tmty{x_n}{A_n}}$
  denotes that the process \tm{P} communicates along channels $\tm{x_1}$, \dots,
  $\tm{x_n}$ following protocols $\ty{A_1}$, \dots, $\ty{A_n}$. Typing
  judgements are derived using rules below.
  \\
  {Structural rules}
  \begin{center}
    \cpInfAx
    \cpInfCut
  \end{center}
	{Logical rules}
  \begin{center}
    \cpInfTens
    \cpInfParr
  \end{center}
  \begin{center}
    \cpInfBot
    \cpInfOne
  \end{center}
  \begin{center}
    \cpInfPlus1
    \cpInfPlus2
  \end{center}
  \begin{center}
    \cpInfWith
  \end{center}
  \begin{center}
    \cpInfNil
    \cpInfTop   
  \end{center}
\end{definition}

\subsection{Metatheory}
\cp enjoys subject reduction, termination, and progress~\citep{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
  \(
    \tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}},
  \)
  such that: no process $\tm{P_i}$ is a cut; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ such that no $\tm{P_i}$ is a cut or a link, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts introduces $n$ channels, and $n+1$ processes. Therefore, at least \emph{one} of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts of $\tm{P}$ such that $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, measured in the size of the cut term. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}

@@

        \hcpEquivMixHalt1

@@

        \hcpEquivMixComm

@@

        \hcpEquivMixAss1

@@

        \hcpEquivNewComm

@@

        \hcpEquivScopeExt1

@@

      \hcpRedAxCut1

@@

      \hcpRedBetaTensParr

@@

      \hcpRedBetaOneBot

@@

      \hcpRedBetaPlusWith1

@@

      \hcpRedBetaPlusWith2

@@

@@

        \SYM{\hcpRedGammaNew}

@@

        \SYM{\hcpRedGammaMix}

@@

        \SYM{\hcpRedGammaEquiv}

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \\
  {Structural rules}
  \begin{center}
    \hcpInfAx
    \hcpInfCut
  \end{center}
  \begin{center}
    \hcpInfMix
    \hcpInfHalt
  \end{center}
  {Logical rules}
  \begin{center}
    \hcpInfBoundTens
    \hcpInfParr
  \end{center}
  \begin{center}
    \hcpInfOne
    \hcpInfBot
  \end{center}
  \begin{center}
    \hcpInfPlus1
    \hcpInfPlus2
  \end{center}
  \begin{center}
    \hcpInfWith
  \end{center}
  \begin{center}
    \hcpInfNil
    \hcpInfTop
  \end{center}
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
\end{definition}\noindent

\subsection{Metatheory}
\hcp enjoys subject reduction, termination, and progress.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
  \(
    \tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}},
  \)
  such that: no process $\tm{P_i}$ is a cut or a mix; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ such that no $\tm{P_i}$ is a cut or a link, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts and mixes introduces $n$ channels. Each application of cut requires an application of mix, so the prefix introduces $n+m+1$ processes. Therefore, at least $\tm{m+1}$ of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts of $\tm{P}$ such that $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}

@@

@@

Disentanglement is terminating, and confluent up to the associativity and commutativity of mixes.
\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If there exists a derivation $\rho$ of $\seq{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp, then there exist derivations $\rho_1, \dots, \rho_n$ of $\seq{\Gamma_1}, \dots, \seq{\Gamma_n}$ in \cp such that
  \[
  \rho
  \moveMixDown^\star
  \begin{prooftree*}
    \AXC{}\noLine\UIC{$\mtf{\rho_1}$}
    \AXC{}\noLine\UIC{$\dots$}
    \AXC{}\noLine\UIC{$\mtf{\rho_n}$}
    \NOM{H-Mix}
    \TIC{$\seq{\Gamma_1 \hsep \dots \hsep \Gamma_n}$}
  \end{prooftree*}
  \]
\end{lemma} 
\begin{proof}
  We repeatedly apply the $\moveMixDown$-rules to the derivation $\rho$ to move the mixes downwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck under a cut, it forms a \cp cut;
  \item
    if a mix gets stuck under a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the top.
  \end{enumerate*}
  All applications of $(\one)$ are followed by an application of \textsc{H-Mix$_0$}, forming a \cp $(\one)$.
\end{proof}
An environment can be internalised as a type by collapsing it as a series of pars.
\begin{definition}\label{def:bigparr}
  \[
  \begin{array}{lcll}
    \ty{\bigparr(\emptycontext)}
    & = & \ty{\bot}
    \\
    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} 
     & = & \ty{A_1 \parr \dots \parr A_n}
     & \text{ if } n \geq 1
    \\
  \end{array}
  \]
\end{definition}\noindent
\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma} 
\begin{proof}
  By repeated application of $(\parr)$.
\end{proof}\noindent
Furthermore, a hyper-environment can be internalised as a type by collapsing it as a series of tensors, where each constituent environment is internalised using $\ty{\bigparr}$. The empty hyper-environment $\ty{\emptyhypercontext}$ is internalised as the unit of tensor.
\begin{definition}\label{def:bigtens}
  \[
  \begin{array}{lcll}
    \ty{\bigtens(\emptyhypercontext)}
    & = & \ty{\one}
    \\
    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
    & \text{ if } n \geq 1
  \end{array}
  \]
\end{definition}\noindent
